---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CAOP.RAA.2024

<!-- badges: start -->
<!-- badges: end -->

The goal of CAOP.RAA.2024 is to provide the official administrative boundaries
of the Azores (Região Autónoma dos Açores (RAA)) as defined in the 2024 edition
of the Carta Administrativa Oficial de Portugal (CAOP), published by the
Direção-Geral do Território (DGT). The package includes convenience functions to
import these boundaries as 'sf' objects for spatial analysis in R.

## Installation

``` r
# install.packages("pak")
pak::pak("patterninstitute/CAOP.RAA.2024")
```

## Introduction

Boundary data is provided for districts (islands), municipalities and civil
parishes. Original source for these data has coordinates projected in UTM 25N for
the western group of islands and UTM 26N for the central and eastern groups.
These projections offer the most accurate planar representation. Use these
functions if you want to use these projections:

- `parishes_25N()` and `parishes_26N()`
- `municipalities_25N()` and `municipalities_26N()`
- `districts_25N()` and `districts_26N()`

Although those functions offer the most accurate projection, they can't be
combined into one single sf object, as this requires the same CRS for all
features. Alternatively, we also offer transformations for EPSG:3035
(ETRS89-extended / LAEA Europe) and EPSG:4326 (WGS 84) which permit having
all islands data in one single sf object. For these data, use these functions:

- `parishes("3035")` or `parishes_26N("4326")`
- `municipalities("3035")` or `municipalities("4326")`
- `districts("3035")` or `districts("4326")`

## Data

Districts (islands):

```{r dev='svg', fig.width=12}
library(CAOP.RAA.2024)
library(ggplot2)

# districts() defaults to districts("3035").
(districts <- districts())
ggplot(mapping = aes(fill = district)) +
  geom_sf(data = districts, col = "white") +
  guides(fill = "none")
```

Municipalities:

```{r dev='svg', fig.width=12}
# municipalities() defaults to municipalities("3035").
(mun <- municipalities())
ggplot(mapping = aes(fill = municipality)) +
  geom_sf(data = mun, col = "white") +
  guides(fill = "none")
```

Parishes of the São Miguel island:

```{r dev='svg', fig.width=12}
# parishes() defaults to parishes("3035").
(parishes <- parishes())
parishes |>
  dplyr::filter(district == "Ilha de São Miguel") |>
  ggplot(mapping = aes(fill = parish)) +
  geom_sf(col = "white") +
  guides(fill = "none")
```

## Sources

- [Carta Administrativa Oficial de Portugal](https://www.dgterritorio.gov.pt/cartografia/cartografia-tematica/caop)
- [GitHub CAOP](https://github.com/dgterritorio/CAOP)
