---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# CAOP.RAA.2024

<!-- badges: start -->
<!-- badges: end -->

The goal of CAOP.RAA.2024 is to provide the official administrative boundaries
of the Azores (Região Autónoma dos Açores (RAA)) as defined in the 2024 edition
of the Carta Administrativa Oficial de Portugal (CAOP), published by the
Direção-Geral do Território (DGT). The package includes convenience functions to
import these boundaries as 'sf' objects for spatial analysis in R.

## Installation

``` r
# install.packages("pak")
pak::pak("patterninstitute/CAOP.RAA.2024")
```

## Data

Districts (islands):

```{r dev='svg', fig.width=12}
library(CAOP.RAA.2024)
library(ggplot2)

(districts <- districts())
ggplot(mapping = aes(fill = district)) +
  geom_sf(data = districts, col = "white") +
  guides(fill = "none")
```

Municipalities:

```{r dev='svg', fig.width=12}
(mun <- municipalities())
ggplot(mapping = aes(fill = municipality)) +
  geom_sf(data = mun, col = "white") +
  guides(fill = "none")
```

Parishes of the São Miguel island:

```{r dev='svg', fig.width=12}
(parishes <- parishes())
parishes |>
  dplyr::filter(district == "Ilha de São Miguel") |>
  ggplot(mapping = aes(fill = parish)) +
  geom_sf(col = "white") +
  guides(fill = "none")
```

## Sources

- [Carta Administrativa Oficial de Portugal](https://www.dgterritorio.gov.pt/cartografia/cartografia-tematica/caop)
- [GitHub CAOP](https://github.com/dgterritorio/CAOP)
